### 2017-02-14 presentation.Rmd

---
title: "Presentation"
author: "Dylan Sun"
date: "2/13/2017"
output: html_document
---

### Comparing treatment groups to control groups
```{r}
t.test(pain[condition == "EDU", .(BPI_PainSeverity_V2)], pain[condition == "CBT", .(BPI_PainSeverity_V2)])
t.test(pain[condition == "EDU", .(BPI_PainSeverity_V2)], pain[condition == "EET", .(BPI_PainSeverity_V2)])

t.test(pain[condition == "EDU", .(BPI_PainSeverity_V3)], pain[condition == "CBT", .(BPI_PainSeverity_V3)])
t.test(pain[condition == "EDU", .(BPI_PainSeverity_V3)], pain[condition == "EET", .(BPI_PainSeverity_V3)])

t.test(pain[condition == "EDU", .(BPI_PainSeverity_V4)], pain[condition == "CBT", .(BPI_PainSeverity_V4)])
t.test(pain[condition == "EDU", .(BPI_PainSeverity_V4)], pain[condition == "EET", .(BPI_PainSeverity_V4)])

```
T-tests comparing the difference in mean pain at baseline found no difference between EDU and EET, and no difference between EDU and CBT.

T-tests comparing the difference in mean pain immediately post-treatment found a significant difference between EDU and EET, and a significant difference between EDU and CBT.

T-tests comparing the difference in mean pain six months post-treatment found no difference between EDU and EET, and no difference between EDU and CBT.